
\documentclass[11pt,twoside,onecolumn,a4paper]{book}

%% put "oui" in the following line to see labels and other debugging text
\newcommand{\imprimeTexteSecret}{non}
%% put "oui" in the following line to see solutions
\newcommand{\imprimeTexteCache}{non}


%%% the following lines controls whether exercises are included
%%% in order to exclude exercises, replace \exostrue by \exosfalse
\newif\ifexos
\exosfalse

\usepackage{lebBook}
\input{lebstuff}
%\input{layoutLNCS.tex}
%\input{layoutOnline.tex}

\UseRawInputEncoding

\usepackage{makeidx}
\makeindex
\input{L10.bdx.tex}
\input{L20.bdx.tex}% this imports the database called  "L20.bdx.tex"
                     % and creates one file per exercise in the temporary
                     % directory defined in lebstuff.tex
                     % this is normally the directory called "temp" in
                     % the father of the directory where the main file
                     % was found by Latex
\input{L21.bdx.tex}
\input{L23.bdx.tex}
\input{L24.bdx.tex}
%\renewtheorem{problem}{Exercise}[chapter]



\begin{document}
%\pagenumbering{arabic}



 \title{
 %\hrulefill\\
 \noindent{\Huge {  NETWORK CALCULUS  }}\\
 ~\\
 A Theory of Deterministic Queuing Systems for the Internet\\
  %\hrulefill\\
 }
 \author{ {\sc Jean-Yves Le Boudec}  \\{\sc Patrick Thiran}\\
 \vspace{3cm}\\
 ~
 Online Version of the Book Springer Verlag - LNCS 2050
 }

% \date{\insfig{smoothingsetting}{0.7}}
\date{Version \today}

 \maketitle
\frontmatter

\begin{titlepage}
  %~\newpage
 \begin{center}
 {\it
 A Annelies\\A Joana, Ma\"{e}lle, Audraine et Elias\\
 A ma m\`{e}re\\
 \hspace{1cm} ---- JL\\
 ~\\
 A mes parents\\
 \hspace{1cm} ---- PT\\
 }
 \vspace{7cm}
 Pour \'{e}viter les grumeaux\\
 Qui encombrent les r\'{e}seaux\\
 Il fallait, c'est compliqu\'{e},\\
 Ma\^{\i}triser les seaux perc\'{e}s\\
 ~\\
  Branle-bas dans les campus \\
  On pourra dor\'{e}navant\\
  Calculer plus simplement\\
 Gr\^{a}ce \`{a} l'alg\`{e}bre Min-Plus\\
 ~\\
  Foin des obscures astuces\\
 Pour estimer les d\'{e}lais\\
 Et la gigue des paquets\\
 Place \`{a} ``Network Calculus"\\
 ~\\
 {\it \hspace{2cm} ---- JL}
 \end{center}
 %\end{flushright}
\setcounter{page}{5}
\end{titlepage}

\textbf{Summary of Changes}
\begin{description}
 \item[2002 Jan 14, JL] Chapter 2: added a better coverage of GR
 nodes, in particular equivalence with service curve. Fixed bug in Proposition 1.4.1
  \item[2002 Jan 16, JL] Chapter 6: M. Andrews brought convincing
  proof that conjecture 6.3.1 is wrong. Redesigned Chapter 6 to
  account for this. Removed redundancy between Section 2.4 and
  Chapter 6. Added SETF to Section 2.4
   \item[2002 Feb 28, JL] Bug fixes in Chapter 9
   \item[2002 July 5, JL] Bug fixes in Chapter 6; changed format for a better printout on most
   usual printers.
   \item[2003 June 13, JL] Added concatenation properties of non-FIFO GR nodes to Chapter
   2. Major upgrade of Chapter 7.
   Reorganized Chapter 7. Added new developments in Diff Serv.
   Added properties of PSRG for non-FIFO nodes.
   \item[2003 June 25, PT] Bug fixes in chapters 4 and 5.
   \item[2003 Sept 16, JL] Fixed bug in proof
   of theorem 1.7.1, proposition 3. The bug was
   discovered and brought to our attention by Fran\c{c}ois Larochelle.
   \item[2004 Jan 7, JL] Bug fix in Proposition 2.4.1 ($\nu >
\frac{1}{h-1}$ instead of $\nu < \frac{1}{h-1}$)
   \item[2004, May 10, JL] Typo fixed in Definition 1.2.4 (thanks
   to Richard Bradford)
   \item[2005, July 13, JL] Bug fixes (thanks
   to Mehmet Harmanci)
   \item[2011, August 17, JL] Bug fixes (thanks
   to Wenchang Zhou)
   \item[2011, Dec 7, JL] Bug fixes (thanks
   to Abbas Eslami Kiasari)
   \item[2012, March 14, JL] Fixed Bug in Theorem 4.4.1
   \item[2012, April 26, JL] Fixed Typo in Section 5.4.2 (thanks to Yuri Osipov)
   \item[2019, March 14, JL] Fixed Typos in Sections 1.4.4 and 2.4.4.
   \item[2019, March 14, JL] Fixed a bug in Proposition 1.3.7. The service curve property is strict only for the high priority flow.
   \item[2019, March 27, JL] Changed the numbering of theorems so that it is the same as with the publisher's version.
   \item[2019, October 1, JL] Fixed typo in Equation (6.2).
   \item[2019, November 26, JL] Fixed typos in Section 1.4.1 ``An ATM Example" (thanks to Hermes Hofmeister Ferreira).

   \item[2019, November 26, JL] Added a comment in Section 1.5.2, ``Example: Buffer Sizing at a Re-shaper" (observe that we
need to make sure that $\sigma(t) = 0$) (triggered by an observation by Hermes Hofmeister Ferreira).
    \item[2019, November 26, JL] Section 1.1.2 clarified that, for left-continuous input and output functions, the delay for a bit that arrives at time $t$ is the limit from the right of $d(t)$, i.e. $d_r(t)\eqdef\limit{s}{t, s>t}d(s)$.
    \item[2019, November 27, JL] Section 1.6.1, Theorem 1.6.3 (minimum delay bound). Modified the statement of the theorem and expanded the proof (triggered by an observation by Hermes Hofmeister Ferreira).
    \item[2019, November 27, JL] Section 1.6.1, Numerical Example after Theorem 1.6.3, corrected bugs (thanks to Hermes Hofmeister Ferreira).
     \item[2019, November 27, JL] Section 1.7.2, Definition 1.7.1, added the condition $\limit{n}{\infty}L(n)=+\infty$ (thanks to Hermes Hofmeister Ferreira).

     \item[2019, November 27, JL] Section 1.7.2, before Equation (1.23) added the condition $r\leq R$ (thanks to Hermes Hofmeister Ferreira).

     \item[2019, November 27, JL] Fixed several typos: in the proofs of Theorem 1.4.7, of Lemma 1.5.1 and of Theorem 1.7.2; and on pages 40 and 45 (thanks to Hermes Hofmeister Ferreira).
     \item[2019, December 9, JL] Added an argument that was implicit in the proof of Theorem 1.7.2.
     \item[2019, December 11, JL] Cosmetic change to the proof of Theorem 1.7.6.
    \item[2019, December 13, JL] Added a missing argument $(t)$ in Definition 3.1.14.
        \item[2020, January 7, JL] Fixed a typo in Section 2.4.3 (thanks to Alaeddine El Fawal).
    \item[2020, September 24, JL] Fixed typos in Theorems 1.6.3; in Theorem 4.3.1, added ``isotone".

    \item[2020, September 25, JL] Fixed typo in Figure 3.3 (thanks to Ludovic Thomas).
    \item[2020, October 12, JL] Fixed typos in: text before Proposition 1.2.1; text before Equation (1.8); proof of Theorem 3.1.11; a few lines after Equation (3.22); last line of proof of Theorem 4.1.4; line before Definition 4.2.2; Table number in Section 4.5; line before Equation (6.8); text between Equations (6.8) and (6.9); notation for max-plus deconvolution in index; text after Definition 7.3.2; Theorem 7.5.1; bullet before Theorem 6.4.1.
        (Thanks to Luxi Zhang.)
        %(Thanks to Zhangjiayi (Jiayi) and the Chinese translators).

        \item[2020, October 13, JL] Replaced the notation $\alpha(t^+)$ by $\alpha_r(t)$ in the proof of Theorem 2.1.4. Changed ``limit to the left" to ``limit from the left". Fixed notation in the proof of Theorem 7.4.5, Item 2. Fixed some formatting issues in the bibliography. (Thanks to Luxi Zhang.)
        \item[2021, June 21, JL] In Section 2.2.5, replaced ``algorithm 3" by ``the algorithm in Definition 2.2.1"


\end{description}
%\end{titlepage}






 \hyphenation{Networks}
\tableofcontents
 \hyphenation{Net-works}

\nfs{\textbf{Version 4}}

\nfs{\textbf{History of Changes}}

\nfs{\textbf{Bug Fixes}}

\nfs{Proposition 1.4.1 corrig\'{e}e}

\nfs{Section 2.4 redesign\'{e}e (diff serv), SETF ajoute, formule
qofis corrig\'{e}e }

\nfs{Section 2.1.3 refaite (GR node)}


\input{L0} % LEB introduction
 \mainmatter
\part{A First Course in Network Calculus}
\input{L20}% LEB: netcal basics
\input{L21}% LEB: appli to Intserv
% remove the following block except for the online version %%%%%%%%%%%%%%%%%%%%
\newpage
\thispagestyle{plain}
\chapter*{~}
\begin{center}


 \noindent{\Huge {  NETWORK CALCULUS  }}\\
 ~\\
 {\Huge {  Parts II and III  }}\\ ~\\
 {\LARGE
 A Theory of Deterministic Queuing Systems for the Internet\\

  %\hrulefill\\

 {\sc Jean-Yves Le Boudec}  \\{\sc Patrick Thiran}\\
 \vspace{3cm}

 Online Version of the Book Springer Verlag - LNCS 2050\\
 Reformatted for improved online viewing and printing\\



% \date{\insfig{smoothingsetting}{0.7}}
 Version \today
 }


\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%% end of block for online version %%%%%%%%%%%%%%%%%
\part{Mathematical Background}

\input{L10}% THI: basic min plus and max plus calculus
\input{L11}% THI: advanced min plus and max plus calculus
%Put exo L30-33 in chapter advanced min plus and max plus calculus
\part{A Second Course in Network Calculus}
\input{L22}% THI: smoothers
\input{L24}% LEB: FIFO
\input{L23}% LEB:   asc + application a Diff serv
\input{L23b}% LEB: time varying
\input{L30}% THI: lossy systems
%\backmatter
\bibliography{leb}
\bibliographystyle{plain}
\printindex
\end{document}
